Supplemental Information: Continuous developmental changes in word recognition support language learning across early childhood

### 2025-10-14 paper/paper.Rmd
correct typos in model specifications (and a couple other small things)

We began by examining developmental changes in children's word recognition. Figure \ref{fig:timecourse} depicts the average timecourse at different ages across all datasets (not controlling for any variation in items and procedures across age groups). Intuitively, these timecourses show gradual increases in accuracy (higher overall proportion target looking) and speed (faster looking to the target after hearing a label) as age increases. To characterize age gradients in speed and accuracy across children, we computed both RTs (reaction times) and accuracies (proportion looking at the target image) following standard practices in the literature [@fernald2008]. Reaction times were computed only on trials for which the child was fixating the distractor at the point of disambiguation (label onset), and were defined as the time from label onset to the first fixation on the target image (see S2 Reaction Times for further details on how reaction times were computed in S2.1 and discussion of issues surrounding distinguishing "correct" vs. "incorrect" trials when computing looking-based reaction times in S2.2).

@@

Our first question was about the functional form of the relationships between age, speed, and accuracy (see S5 Pairwise Correlations of Main Measures for raw pairwise correlations between variables). We began by fitting linear mixed-effects models predicting speed and accuracy on each trial across the full dataset with random slopes of child age nested within study (modeling item and procedural variation across studies) and random intercepts by participant. We compared models that included both long and short accuracy windows, as well as logarithmic and linear effects of age, and logarithmic and linear transformations of RT (see S3 Checks on Data Distributional Assumptions for further analyses and discussion of these modeling choices). The best fitting model of accuracy predicted long window accuracy as a function of the logarithm of age; the best fitting model of speed predicted log RT as a function of log age as well (see S6 Functional Form Model Comparison and S7 Power Law Fits). Because long window accuracies were more correlated with other variables and showed clearer age gradients, we focus on these in our analyses.

>>>>

We began by examining developmental changes in children's word recognition. Figure \ref{fig:timecourse} depicts the average timecourse at different ages across all datasets (not controlling for any variation in items and procedures across age groups). Intuitively, these timecourses show gradual increases in accuracy (higher overall proportion target looking) and speed (faster looking to the target after hearing a label) as age increases. To characterize age gradients in speed and accuracy across children, we computed both RTs (reaction times) and accuracies (proportion looking at the target image) following standard practices in the literature [@fernald2008]. Reaction times were computed only on trials for which the child was fixating the distractor at the point of disambiguation (label onset), and were defined as the time from label onset to the first fixation on the target image (see S2 Reaction Times, including further details on how reaction times were computed in S2.1 and discussion of issues surrounding distinguishing "correct" vs. "incorrect" trials when computing looking-based reaction times in S2.2).

@@

Our first question was about the functional form of the relationships between age, speed, and accuracy (see S5 Pairwise Correlations of Main Measures for raw pairwise correlations between variables). We began by fitting linear mixed-effects models predicting speed and accuracy on each trial across the full dataset with random slopes of child age nested within study (modeling item and procedural variation across studies) and random intercepts by participant (see S8 Mixed-effects model specifications for further details on these specifications). We compared models that included both long and short accuracy windows, as well as logarithmic and linear effects of age, and logarithmic and linear transformations of RT (see S3 Checks on Data Distributional Assumptions for further analyses and discussion of these modeling choices). The best fitting model of accuracy predicted long window accuracy as a function of the logarithm of age; the best fitting model of speed predicted log RT as a function of log age as well (see S6 Functional Form Model Comparison and S7 Power Law Fits). Because long window accuracies were more correlated with other variables and showed clearer age gradients, we focus on these in our analyses.

### 2025-10-14 paper/supplemental_information.Rmd
correct typos in model specifications (and a couple other small things)

```{r rt-hist, fig.cap="\\label{fig:rt-hist} Distribution of RT overlaid with log normal distribution, split by age.", fig.pos="h!", fig.width=8, fig.height=5}

@@

Table \ref{tab:accmods} shows model comparison measures between for different models of the functional form of the relationship between accuracy and age and Table \ref{tab:rtmods} shows the same for reaction time. Age gradients are estimated substantially better with long window accuracies. Note that there are a greater number of observations for short window accuracies due to less missing data. We speculate that, on average, more participants looked away from the screen towards the end of trials, leading to a greater number of exclusions of long window trials based on the 50% criterion. Note that the total percentage of trials excluded is still small for both measures: `r round(mean(is.na(d_trial$long_window_accuracy)), 3)*100`% for long window accuracy and `r round(mean(is.na(d_trial$short_window_accuracy)), 3)*100`% for short window accuracy.

@@

With our current dataset we cannot specifically determine whether within-individual patterns of change conform to linear, power law, or exponential developmental patterns unfortunately, because we have insufficient data about individuals' improvement across time. Thus, our current results apply to the form of the age gradient as opposed to the form of any individual's pattern of developmental change.

@@

\begin{verbatim}
long_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
long_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
short_window_accuracy ~ age_s  + (age_s | dataset_name) + (1 | subject_id)
short_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
\end{verbatim}

>>>>

```{r rt-hist, fig.cap="\\label{fig:rt-hist} Distribution of RT overlaid with a log normal distribution, split by age.", fig.pos="h!", fig.width=8, fig.height=5}

@@

Table \ref{tab:accmods} shows model comparison measures for different models of the functional form of the relationship between accuracy and age and Table \ref{tab:rtmods} shows the same for reaction time. Age gradients are estimated substantially better with long window accuracies. Note that there are a greater number of observations for short window accuracies due to less missing data. We speculate that, on average, more participants looked away from the screen towards the end of trials, leading to a greater number of exclusions of long window trials based on the 50% criterion. Note that the total percentage of trials excluded is still small for both measures: `r round(mean(is.na(d_trial$long_window_accuracy)), 3)*100`% for long window accuracy and `r round(mean(is.na(d_trial$short_window_accuracy)), 3)*100`% for short window accuracy.

@@

With our current dataset, we unfortunately cannot specifically determine whether within-individual patterns of change conform to linear, power law, or exponential developmental patterns, because we have insufficient data about individuals' improvement across time. Thus, our current results apply to the form of the age gradient as opposed to the form of any individual's pattern of developmental change.

@@

\begin{verbatim}
long_window_accuracy ~ age_s + (age_s | dataset_name) + (1 | subject_id)
long_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
short_window_accuracy ~ age_s  + (age_s | dataset_name) + (1 | subject_id)
short_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
\end{verbatim}
